Implications of uniformly distributed, empirically informed priors for phylogeographical model selection: A reply to Hickerson et al.

### 2014-07-02 manuscripts/reply/reply-main.tex
Working on reply ms.
Moving sampling error figure and the plot of Hickerson et al.'s
mixture-model prior on divergence times to the Supporting Information.

To explore this prediction further, we repeat the analysis of the Philippines
dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim
U(0, 10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{}
\sim (0.0005, 0.02)$), using a very large prior sample size of $10^8$.
When we look at the trace of the estimates of the dispersion index of
divergence times (\vmratio{}) as the prior samples accumulate
(Figure~\ref{figSamplingError}) we do not see the trend predicted by
\citet{Hickerson2013} in either the unadjusted or GLM-regression-adjusted
estimates.
While sampling error is always a reality, it does not appear to be playing a
large role in the biases revealed by the results of \citet{Oaks2012} or
presented above.

@@

The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
of \msb are similar to those of \citet{Hickerson2013}, but avoid the
need for imposing an additional dimension of model choice.
\citet{Oaks2012} suggest that uniform priors may not be appropriate for many
parameters of the \msb model, and recommend the use of probability
distributions from the exponential family.
If we look at the prior distribution on divergence-time parameters imposed by
the model-averaging approach of \citet{Hickerson2013} we see it is a mixture of
overlapping uniforms with lower limits of zero (Figure~\ref{figMCTauPrior}).
This looks very much like an exponential distribution, except that in any state
of the model, all the divergence-time parameters are restricted to the hard
bounds of one of the uniform distributions.
Thus, it seems more appropriate to simply place a gamma prior (the exponential
being a special case of the gamma) on divergence times.
This would capture the prior uncertainty that \citet{Hickerson2013} are
suggesting for divergence times (Figure~\ref{figMCTauPrior}) while avoiding
costly model-averaging and the constraint that all divergence times must fall
within the hard bounds of the current model state.
It also would allow an investigator to place the majority of the prior density
in regions of parameter space they believe, \emph{a priori}, are most
plausible, but still capture uncertainty in the tails of distributions with low
density.
Most importantly, we have found the use of more flexible distributions in place
of uniform priors improves the power of the method to detect temporal variation
in divergences and reduces the bias toward models of clustered divergences
\citep{Oaks2014dpp}.

>>>>

To explore this prediction further, we repeat the analysis of the Philippines
dataset under the intermediate prior used by \citet{Oaks2012} ($\divt{} \sim
U(0, 10)$, $\meanDescendantTheta{} \sim (0.0005, 0.04)$, $\ancestralTheta{}
\sim (0.0005, 0.02)$), using a very large prior sample size of $10^8$.
When we look at the trace of the estimates of the dispersion index of
divergence times (\vmratio{}) as the prior samples accumulate
(Figure~S\ref{figSamplingError}) we do not see the trend predicted by
\citet{Hickerson2013} in either the unadjusted or GLM-regression-adjusted
estimates.
While sampling error is always a reality, it does not appear to be playing a
large role in the biases revealed by the results of \citet{Oaks2012} or
presented above.

@@

The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
of \msb are similar to those of \citet{Hickerson2013}, but avoid the
need for imposing an additional dimension of model choice.
\citet{Oaks2012} suggest that uniform priors may not be appropriate for many
parameters of the \msb model, and recommend the use of probability
distributions from the exponential family.
If we look at the prior distribution on divergence-time parameters imposed by
the model-averaging approach of \citet{Hickerson2013} we see it is a mixture of
overlapping uniforms with lower limits of zero (Figure~S\ref{figMCTauPrior}).
This looks very much like an exponential distribution, except that in any state
of the model, all the divergence-time parameters are restricted to the hard
bounds of one of the uniform distributions.
Thus, it seems more appropriate to simply place a gamma prior (the exponential
being a special case of the gamma) on divergence times.
This would capture the prior uncertainty that \citet{Hickerson2013} are
suggesting for divergence times (Figure~S\ref{figMCTauPrior}) while avoiding
costly model-averaging and the constraint that all divergence times must fall
within the hard bounds of the current model state.
It also would allow an investigator to place the majority of the prior density
in regions of parameter space they believe, \emph{a priori}, are most
plausible, but still capture uncertainty in the tails of distributions with low
density.
Most importantly, we have found the use of more flexible distributions in place
of uniform priors improves the power of the method to detect temporal variation
in divergences and reduces the bias toward models of clustered divergences
\citep{Oaks2014dpp}.

### 2014-07-02 manuscripts/reply/reply-si-figures.tex
Updating reply ms.
Moving coin flip figure to SI, changing to right justified, minor edit
to section title.
Working on reply ms.
Moving sampling error figure and the plot of Hickerson et al.'s
mixture-model prior on divergence times to the Supporting Information.

@@

@@

>>>>

\siFigure{../images/coin_flip.pdf}{
    A plot of three beta probability density functions that represent a prior
    (black; $beta(10, 10)$), posterior (blue; $beta(13, 17)$), and empirical
    Bayes density (red; $beta(16, 24)$) for a dataset of 10 coin flips, three
    of which are successes.
}{figCoinFlip}

@@

%% Sampling error
\widthFigure{0.9}{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{listformat=sFigList}{
    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
    from the prior. Both (A) unadjusted and (B) GLM-regression-adjusted estimates are shown.
    The data analyzed were the 22 pairs of Philippine taxa from \citet{Oaks2012}.
    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
}{figSamplingError}
\clearpage

@@

%% Model-choice tau prior
\siFigure{../../model-choice/tau-prior/tau_prior.pdf}{
    The prior distribution on divergence times imposed by the model-averaging prior
    comprised of five models with different uniform priors on \divt{}:
    $M_1$ ($\divt{} \sim U(0, 0.1)$), $M_2$ ($\divt{} \sim U(0, 1)$), $M_3$
    ($\divt{} \sim U(0, 5)$), $M_4$ ($\divt{} \sim U(0, 10)$), $M_5$ ($\divt{}
    \sim U(0, 20)$).
}{figMCTauPrior}
